Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.Rmd
update 2.2.1 text

<!-- improve the first sentence -->

One may prefer conventional hypothesis testing because reading diagnostic plots is to some extent subjective and indecisive as it relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984]. It is arguably true that visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data. One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].

@@

<!-- use true exp lineup as example lineup -->

```{r first-example-lineup, fig.cap="Visual testing is conducted using a lineup, as in the example here. The residual plot computed from the observed data (plot 11, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model. Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot. A small $p$-value would result from a substantial number selecting plot 11."}
mod <- poly_model(include_z = TRUE, sigma = 1)
POLY_MODEL$plot_lineup(mod$gen_lineup(300, pos = 11), theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
```

Figure \ref{fig:first-example-lineup} is an example of a lineup protocol. If the actual data plot at position 11 is identifiable, then it is evidence for the rejection of the null hypothesis that the regression model is correctly specified. In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity issue.

The effectiveness of lineup protocol has already been validated by @majumder_validation_2013 under relatively simple classical normal linear regression model settings with only one or two regressors. Their results suggest visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they expressed that in general it is unnecessary to use visual inference if there exists a conventional test and they didn't expect the visual test to perform equally well as the conventional test. In their third experiment, where there does not exist a proper conventional test, visual test outperforms the conventional test for a large margin. This is encouraging as it promotes the use of visual inference in border field of data science where there are no existing statistical testing procedures. In fact, lineup protocol has been integrated successfully into diagnostic tools of hierarchical linear models [@loy2013diagnostic].

<!-- adam work -->

>>>>

One may argue that reading diagnostic plots is to some extent subjective and indecisive compared to those rigorous statistical procedures as it relies on graphical perception - human's ability to interpret and decode the information embedded in the graph [@cleveland_graphical_1984]. It is particularly true that visual discovery suffers from its unsecured and unconfirmed nature where the degree of the presence of the visual features typically can not be measured quantitatively and objectively, which may lead to over or under-interpretations of the data. One such example is finding an over-interpretation of the separation between gene groups in a two-dimensional projection from a linear discriminant analysis when in fact there are no differences in the expression levels between the gene groups and separation is not an uncommon occurrence [@roy_chowdhury_using_2015].

@@

```{r first-example-lineup, fig.cap="Visual testing is conducted using a lineup, as in the example here. The residual plot computed from the observed data (plot 7, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model. Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot. A small $p$-value would result from a substantial number selecting plot 7."}

vi_lineup$poly_422$data %>%
  VI_MODEL$plot_lineup(theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
```

Figure \ref{fig:first-example-lineup} is an example of a lineup protocol. If the actual data plot at position 7 is identifiable, then it is evidence for the rejection of the null hypothesis that the regression model is correctly specified. In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity issue.

The effectiveness of lineup protocol is validated by @majumder_validation_2013 under relatively simple classical normal linear regression model settings with one or two regressors. Their results suggest visual test is capable of testing the significance of a single regressor with a similar power as a t-test, though they expressed that in general it is unnecessary to use visual inference if there exists a conventional test and they didn't expect the visual test to perform equally well as the conventional test. In their third experiment, where there does not exist a proper conventional test, visual test outperforms the conventional test for a large margin. This is encouraging as it promotes the use of visual inference in border field of data science where there are no existing statistical testing procedures. Visual inference have also been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you. They use lineup protocols to judge the assumption of linearity, normality and constant error variance for both the level-1 and level-2 residuals.

### 2022-12-08 paper_comparison/paper_comparison/paper_comparison.tex
update 2.2.1 text

Unlike hypothesis testing built upon rigorous statistical procedures,
reading diagnostic plots relies on graphical perception - human's
ability to interpret and decode the information embedded in the graph
\citep{cleveland_graphical_1984}, which is to some extent subjective and
indecisive. Further, visual discovery suffers from its unsecured and
unconfirmed nature where the degree of the presence of the visual
features typically can not be measured quantitatively and objectively,
which may lead to over or under-interpretations of the data. One such
example is finding an over-interpretation of the separation between gene
groups in a two-dimensional projection from a linear discriminant
analysis when in fact there are no differences in the expression levels
between the gene groups and separation is not an uncommon occurrence
\citep{roy_chowdhury_using_2015}.

@@

\caption{Visual testing is conducted using a lineup, as in the example here. The residual plot computed from the observed data (plot 11, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model. Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot. A small $p$-value would result from a substantial number selecting plot 11.}\label{fig:first-example-lineup}
\end{figure}

Figure \ref{fig:first-example-lineup} is an example of a lineup
protocol. If the actual data plot at position 11 is identifiable, then
it is evidence for the rejection of the null hypothesis that the
regression model is correctly specified. In fact, the actual residual
plot is obtained from a misspecified regression model with non-linearity
issue.

The effectiveness of lineup protocol has already been validated by
\citet{majumder_validation_2013} under relatively simple classical
normal linear regression model settings with only one or two regressors.
Their results suggest visual test is capable of testing the significance
of a single regressor with a similar power as a t-test, though they
expressed that in general it is unnecessary to use visual inference if
there exists a conventional test and they didn't expect the visual test
to perform equally well as the conventional test. In their third
experiment, where there does not exist a proper conventional test,
visual test outperforms the conventional test for a large margin. This
is encouraging as it promotes the use of visual inference in border
field of data science where there are no existing statistical testing
procedures. In fact, lineup protocol has been integrated successfully
into diagnostic tools of hierarchical linear models
\citep{loy2013diagnostic}.

>>>>

One may argue that reading diagnostic plots is to some extent subjective
and indecisive compared to those rigorous statistical procedures as it
relies on graphical perception - human's ability to interpret and decode
the information embedded in the graph \citep{cleveland_graphical_1984}.
It is particularly true that visual discovery suffers from its unsecured
and unconfirmed nature where the degree of the presence of the visual
features typically can not be measured quantitatively and objectively,
which may lead to over or under-interpretations of the data. One such
example is finding an over-interpretation of the separation between gene
groups in a two-dimensional projection from a linear discriminant
analysis when in fact there are no differences in the expression levels
between the gene groups and separation is not an uncommon occurrence
\citep{roy_chowdhury_using_2015}.

@@

\caption{Visual testing is conducted using a lineup, as in the example here. The residual plot computed from the observed data (plot 7, exhibiting non-linearity) is embedded among 19 null plots, where the residuals were simulated from a standard error model. Computing the $p$-value requires that the lineup be examined by a number of human judges, each asked to select the most different plot. A small $p$-value would result from a substantial number selecting plot 7.}\label{fig:first-example-lineup}
\end{figure}

Figure \ref{fig:first-example-lineup} is an example of a lineup
protocol. If the actual data plot at position 7 is identifiable, then it
is evidence for the rejection of the null hypothesis that the regression
model is correctly specified. In fact, the actual residual plot is
obtained from a misspecified regression model with non-linearity issue.

The effectiveness of lineup protocol is validated by
\citet{majumder_validation_2013} under relatively simple classical
normal linear regression model settings with one or two regressors.
Their results suggest visual test is capable of testing the significance
of a single regressor with a similar power as a t-test, though they
expressed that in general it is unnecessary to use visual inference if
there exists a conventional test and they didn't expect the visual test
to perform equally well as the conventional test. In their third
experiment, where there does not exist a proper conventional test,
visual test outperforms the conventional test for a large margin. This
is encouraging as it promotes the use of visual inference in border
field of data science where there are no existing statistical testing
procedures. Visual inference have also been integrated into diagnostic
of hierarchical linear models by \citet{loy2013diagnostic},
\citet{loy2014hlmdiag} and \citet{loy2015you}. They use lineup protocols
to judge the assumption of linearity, normality and constant error
variance for both the level-1 and level-2 residuals.
